
\section{Introduction}

Public SPARQL endpoints cannot fully execute many SPARQL queries due
to their quotas and fair-use policies. After 60 seconds, they stop
their execution and return only partial results, forbidding many use
cases such as building summaries or computing large-scale statistics.
This constitutes a major issue for building decentralized ecosystems
of public SPARQL endpoints.

Sampling-based Approximate Query Processing
(S-AQP)~\cite{DBLP:conf/sigmod/AgarwalMKTJMMS14} tackles this issue
for use cases such as computing large-scale
statistics~\cite{soulet2019anytime,10.1007/978-3-319-18818-8_14},
knowledge graph embeddings~\cite{ristoski2016rdf2vec}, join
orders~\cite{DBLP:conf/cidr/LeisRGK017}, approximate
aggregations~\cite{wang2022approximate},
summaries~\cite{10.1007/978-3-030-49461-2_10}, and exploratory
queries~\cite{DBLP:conf/sigmod/AgarwalMKTJMMS14}.  By confining query
execution to samples of large datasets, S-AQP drastically reduces
execution time, delivering approximate results with error estimates.
%
To support S-AQP in SPARQL, the evaluation of such queries \emph{must}
both
\begin{inparaenum}[(i)]
\item return random samples, and
\item comply with fair-use policies of public SPARQL endpoints.
\end{inparaenum}
Ad-hoc methods already exist to sample knowledge graphs.  For example,
a user could draw random triples from
Wikidata~\cite{soulet2019anytime} by repeatedly executing
\verb|SELECT * {?s ?p ?o} OFFSET r LIMIT 1|, where $r$ is a random
number between $0$ and the dataset size ($0<r<12B$). However, all
queries time out when $r$ is above $100M$.
%Hala: random sampling from the set of results matched by a particular SPARQL triple pattern
Some triple stores propose home-made methods for sampling triple
patterns%
\footnote{\url{https://docs.stardog.com/query-stardog/sampling-service\#sampling-service}}\textsuperscript{,}\footnote{\url{https://docs.openlinksw.com/virtuoso/rndsalltr/}},
but these solutions are limited to single triple patterns and the
underlying complexity is not established.

In this demonstration, we introduce \NAME, an extension of Apache Jena
that efficiently supports S-AQP for conjunctive SPARQL queries. \NAME
evaluates a query $Q$ over a dataset $D$ by drawing random walks
guided by $Q$, as defined in \WANDER~\cite{li2019wanderjoin}. Random
walks have two desirable properties:
\begin{inparaenum}[(i)]
\item As each random walk is \emph{independant}, a user can distribute
  its query execution into multiple requests, hence collecting and
  merging random walks in a pay-as-you-go fashion without impairing
  the fair-use policy of public SPARQL endpoints.
\item As each random walk enjoys a \emph{logarithmic time complexity},
  endpoints can draw thousands of random walks per second: the widely
  used BTree indexes of RDF stores allow endpoints to compute each
  random walk in $\mathcal{O}(|Q| \log |D|)$ where $|Q|$ is the number
  of triple/quad patterns in $Q$ and $|D|$ is the number of
  triples/quad in $D$.
\end{inparaenum}

\NAME produces random partial results instead of complete
results. Without cardinality estimates, users cannot appreciate the
quality of provided samples. In Figure~\ref{fig:raw_screenshot}, \NAME
returns $1434$ random results after $35$ seconds. By adding an
estimate of the total number of results ($26 \pm 3M$), the user
acknwoledges that her sample represents but a tiny fraction of the
whole results.  Provided stimates prove highly accurate as long as
random walks remain paired with their probability of being
drawn~\cite{DBLP:conf/sigmod/ParkKBKHH20}. Despite starting from $35
\pm 15M$, \NAME improved over time, reaching $26\pm 3M$ while the
actual number of results is $25M$.

For the demonstration, we load the dump of Wikidata provided by
WDBench~\cite{angles2022wdbench} that comprises $1.2B$ triples.  We
let users choose a query among the $16$ that time out on Apache
Jena. We execute it using \NAME to highlight
\begin{inparaenum}[(i)]
\item the benefits of S-AQP for ecosystems of public SPARQL endpoints,
  and
\item the feasibility of implementing S-AQP on well-known engines.
\end{inparaenum}

%% the $1.2B$ triples Wikidata dataset
%% provided by WDBench~\cite{angles2022wdbench}, and we re-execute the 16
%% queries that time out on Jena using \NAME. %S-AQP as explained in
%% Section~\ref{proposal}.
%
% Hala
%  The S-AQP results shed light on the reasons for these queries time out and provide users with valuable  information about
% Thanks to S-AQP results, users can understand
% why these queries time out. Overall, this demonstration highlights


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper.tex"
%%% End:
